---
title: "Advent of Code 2017"
subtitle: "Day 1"
format: html
---

## Shifting and comparing

### Part 1

Having a sequence of digits, one digit with the next (or, if the digit is the last, the first digit in the sequence).
The digits that are the same as the next digit, are added together. Do that for a very large number.

#### The example

- 1122 produces a sum of 3 (1 + 2) because the first digit (1) matches the second digit and the third digit (2) matches the fourth digit.
- 1111 produces 4 because each digit (all 1) matches the next.
- 1234 produces 0 because no digit matches the next.
- 91212129 produces 9 because the only digit that matches the next one is the last digit, 9.

Construct the test-cases:
```{r part-1-example}
example <- c("1122", "1111", "1234", "91212129")
```

Why a character vector? The input for the actual task might technically be a number. But it is so huge
that we do not want to handle it as such.

We split the input, and convert to a numeric vector. Next that vector is shiftet to the right,
placing the last value as the first, the first as the second etc. Then we compare those two vectors,
and use that as a boolean index to the values. 
```{r}
library(tidyverse)

captcha <- function(input){
  input <- str_split_1(input, "") |> as.numeric()
  input_shifted <- c(input[length(input)], input[2:length(input)-1])
  sum(input[input == input_shifted])
}

```

Testning:

```{r}
captcha(example[4])
```

It works.

#### The actual task

Read in the input:
```{r}
input <- readLines("data/day-01-input.txt")
```

Use it as input to the function:

```{r}
captcha(input)
```

First star!

### Part 2

Next we need to shift the vector even more. 50% in fact. This requires a new captcha function:

```{r part-2-function}
next_captcha <- function(input){
    input <- str_split_1(input, "") |> as.numeric()
    len <- length(input)
    input_shifted <- c(input[(len/2+1):len], input[1:(len/2)])
    sum(input[input == input_shifted])
  }
```

#### The example

- 1212 produces 6: the list contains 4 items, and all four digits match the digit 2 items ahead.
- 1221 produces 0, because every comparison is between a 1 and a 2.
- 123425 produces 4, because both 2s match each other, but no other digit has a match.
- 123123 produces 12.
- 12131415 produces 4.

Setting up example data:

```{r}
example <- c("1212", "1221", "123425", "123123", "12131415")
```

Testing:

```{r}
next_captcha(example[2])
```

#### The actual task

```{r}
next_captcha(input)
```

Done 🎄