---
title: "aoc 2025 - dag 1"
format: html
---

If we treat the position of the dial as a number modulo 100, we can simply
add or subtract the values, and count the number of times the cumulative result
is 0.

Libraries:
```{r}
library(tidyverse)
```

The example data:

```{r}
example <- c("L68",
"L30",
"R48",
"L5",
"R60",
"L55",
"L1",
"L99",
"R14",
"L82")
```

Enter it into at tibble, separate the input into two columns, one for the direction, and one for the value. Transform direction to a signed value, calculate the cumultative sum, and add 50 which is the starting postion:

```{r}
sequence <- tibble(input = example) |> 
  separate(input, into = c("dir", "value"), sep = "(?<=[A-Za-z])(?=[0-9])") |> 
  mutate(value = as.numeric(value)) |> 
  mutate(value = case_when(
    dir == "R" ~ value,
    dir == "L" ~ - value
  )) |> 
  mutate(cumsum = cumsum(value)) |> 
  mutate(cumsum = cumsum +50)
```

Take modulus 100 to the values, are they 0? Get the sum, theres the answer (to the example data):
```{r}
sum(sequence$cumsum %% 100 == 0)
```

Now on the real data:

```{r}
sequence <- tibble(input = readLines("aoc2025/input-01.txt")) |> 
  separate(input, into = c("dir", "value"), sep = "(?<=[A-Za-z])(?=[0-9])") |> 
  mutate(value = as.numeric(value)) |> 
  mutate(value = case_when(
    dir == "R" ~ value,
    dir == "L" ~ - value
  )) |> 
  mutate(cumsum = cumsum(value)) |> 
  mutate(cumsum = cumsum +50)
sum(sequence$cumsum %% 100 == 0)
```

Part one done!

## Part two

Now we need to figure out how many times the pointer passes 0. So rather than calculating the position after each step, we take it in steps of one.

Same principle as before, but now with 50 individual steps to the left, rather than a single 50 step long jump to the left.
```{r}
sequence <-  tibble(input = example) |> 
  separate(input, into = c("dir", "value"), sep = "(?<=[A-Za-z])(?=[0-9])") |> 
  mutate(value = as.numeric(value)) |> 
  mutate(value = case_when(
    dir == "R" ~ value,
    dir == "L" ~ -value
  )) |> 
   mutate(test = map(value, ~sign(.x)*rep(1, times = abs(.x)))) |> 
   unnest(test) |> 
  mutate(cumsum = cumsum(test)) |> 
  mutate(cumsum = cumsum +50)

sum(sequence$cumsum %% 100 == 0)
```

Correct test result, now the complete task:

```{r}
sequence <-  tibble(input = readLines("aoc2025/input-01.txt")) |> 
  separate(input, into = c("dir", "value"), sep = "(?<=[A-Za-z])(?=[0-9])") |> 
  mutate(value = as.numeric(value)) |> 
  mutate(value = case_when(
    dir == "R" ~ value,
    dir == "L" ~ -value
  )) |> 
   mutate(test = map(value, ~sign(.x)*rep(1, times = abs(.x)))) |> 
   unnest(test) |> 
  mutate(cumsum = cumsum(test)) |> 
  mutate(cumsum = cumsum +50)

sum(sequence$cumsum %% 100 == 0)
```

Done.