---
title: "Advent of Code 2016"
subtitle: "Day "
format: html
---

## Lookup toilets

A tale of two lookup tables.

```{r lib-load}
library(stringr)
```

### Part 1

On a keypad, start somewhere, follow instructions moving around. First set of 
instructions give the first value. Second set of instructions, starting at the 
first value from the previous set of instructions, give the second value.
Continue until you run out of instructions.

#### The example

This the the keypad. We start at 5, instructions "UU" moves up to "2". The second "U" does not
change anything, because we cannot go up from that

123
456
789

We make a lookup table. Given a position and an instruction, what is the new position?

```{r}
lut <- c( 1, 4, 1, 2,
 2, 5, 1, 3,
 3, 6, 2, 3,
 1, 7, 4, 5,
 2, 8, 4, 6,
 3, 9, 5, 6,
 4, 7, 7, 8,
 5, 8, 7, 9,
 6, 9, 8, 9) |> 
  matrix(nrow = 9, byrow = T,
    dimnames = list(pos = c(1:9), instruction = c("U", "D", "L", "R")))
lut
```

And a function for updating the position:

```{r}
new_pos <- function(pos, instruction){
  lut[pos, instruction]
}
```

Construct the instruction set:

```{r}
input <- c("ULL", "RRDDD", "LURDL", "UUUUD")
```

Split and save in list:

```{r}
input <-   str_split(input, "")
```

Write a function that loops through the length of the instruction set,
and in an inner loop loops through the instructions in each element.
```{r}
answer <- numeric()
pos <- 5
for(i in 1:length(input)){
  for(j in input[[i]]){
    pos <- new_pos(pos, j)
  }
answer[i] <- pos
}

answer <- paste0(answer, collapse = "")
```

Returns the correct answer!

#### The actual task

Now read in the complete instruction set:

```{r}
input <- readLines("aoc2016/data/02-input.txt")
input <-   str_split(input, "")

answer <- numeric()
pos <- 5
for(i in 1:length(input)){
  for(j in input[[i]]){
    pos <- new_pos(pos, j)
  }
answer[i] <- pos
}

answer <- paste0(answer, collapse = "")

```

### Part 2

Basically the exact same task. But with different layout of keypad.



#### The example

New lookup-table:

```{r}
lut <- c("1",	"3",	"1",	"1",
"2",	"6",	"2",	"3",
"1",	"7",	"2",	"4",
"4",	"8",	"3",	"4",
"5",	"5",	"5",	"6",
"2",	"A",	"5",	"7",
"3",	"B",	"6",	"8",
"4",	"C",	"7",	"9",
"9",	"9",	"8",	"9",
"6",	"A",	"A",	"B",
"7",	"D",	"A",	"C",
"8",	"C",	"B",	"C",
"B",	"D",	"D",	"D") |> 
    matrix(nrow = 13, byrow = T,
    dimnames = list(pos = c(as.character(1:9), LETTERS[1:4]), instruks = c("U", "D", "L", "R")))

lut
```

The test input:

```{r}
input <- c("ULL", "RRDDD", "LURDL", "UUUUD")
input <-   str_split(input, "")
```

And apply the same procedure. But remember that we now have characters on
the keypad.

```{r}
answer <- character()
pos <- "5"
for(i in 1:length(input)){
  for(j in input[[i]]){
    pos <- new_pos(pos, j)
  }
answer[i] <- pos
}

answer <- paste0(answer, collapse = "")
answer
```

Nice

#### The actual task

The exact same procedure, now with the full input:

```{r}
input <- readLines("aoc2016/data/02-input.txt")
input <-   str_split(input, "")


answer <- character()
pos <- "5"
for(i in 1:length(input)){
  for(j in input[[i]]){
    pos <- new_pos(pos, j)
  }
answer[i] <- pos
}

answer <- paste0(answer, collapse = "")

```

Done 🎄